Biodiversity temporal trends are reshaping food web structure and redundancy in riverine ecosystems

### 2025-04-10 paper/manuscript.Rmd
Review abstract, Significance statement and appendix 1

Understanding how biodiversity loss affects ecological communities is crucial
for predicting and mitigating human impacts on ecosystems. This study provides
comprehensive evidence on how biodiversity changes are reshaping food web
structure using an extensive dataset of riverine fish communities. We
demonstrate that temporal changes in species richness and community biomass
consistently correlate with altered trophic levels, connectance, and the
redundancy of feeding relationships, suggesting energetic constraints and
biodiversity-ecosystem functioning mechanisms jointly shape food web assembly
and disassembly. Importantly, we show that declining biodiversity reduces the
redundancy of trophic pathways, potentially making communities more vulnerable
to future perturbations. These results suggest that food web structure metrics
could serve as valuable indicators for monitoring ecosystem health and
evaluating conservation strategies.

>>>>

Current rates of biodiversity loss raise urgent concerns about ecosystem
functioning and stability, yet how biodiversity trends directly impact these
critical processes remains unclear. For instance, food web &
Biodiversity-Ecosystem Functioning research delivered strong predictions
about the ecological consequences of biodiversity loss. Using an extensive
dataset of riverine fish communities, we link biodiversity trends to temporal
changes in food web structure. We demonstrate that temporal changes in species
richness and community biomass consistently associate with altered trophic
levels, connectance, and the redundancy of feeding relationships. We found that
declining biodiversity reduces trophic pathway redundancy, likely increasing
ecosystem vulnerability to perturbations. Our findings suggest that food web
structure metrics could serve as valuable indicators for monitoring ecosystem
health and guiding conservation strategies.
